\documentclass{article}
\usepackage{hyperref}
\usepackage{color}
\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in}
\addtolength{\textwidth}{1.75in}
\addtolength{\topmargin}{-.875in}
\addtolength{\textheight}{1.75in}
\pagestyle{plain}
\author{Rahul Daware}
\begin{document}
\title{Chapter 5 : Elastic Compute Cloud (EC2)}
\maketitle
\newpage
\tableofcontents
\newpage

Amazon Elastic Compute Cloud (Amazon EC2) is a web service that provides computer capacity in the cloud. Amazon EC2 reduces the time required to obtain an boot new server instances to minutes, allowing you to quickly scale capacity, both up and down, as your computing requirements change.

Aamzon EC2 changes the economics of computing by allowing you to pay only for capacity that you will actually use. Amazon EC2 provides developers the tools to build failure resilient applications and isolate themselves from common failure scenarios.

\section{EC2 Options}
\begin{itemize}
\item
On Demand - allow you to pay a fixed rate by the hour (or by the second) with no commitment

\item
Reserved - Provide yo with a capacity reservation, and offer a significant discount on the hourly charge for an instance. 1 Year or 3 Year terms

\item
Spot - Enables you to bid whatever price you want for instance capacity,providing for even greater savings if your applications have flexible start and end times

\item
Dedicated Hosts - Physical EC2 server dedicated for your use. Dedicated hosts can help you reduce costs by allowing you to use your existing server-bound software licenses
\end{itemize}

\subsection{On Demand Instances}
\begin{itemize}
\item
Users that want the low cost and flexibility of Amazon EC2 without any upfront payment or long term commitment

\item
Applications with short term, spiky, or unpredictable workloads that cannot be interrupted

\item
Applications being developed or tested on Amazon EC2 for the first time
\end{itemize}

\subsection{Reserved Instances}
\begin{itemize}
\item
Applications with steady state or predictable usage

\item
Applications that require reserved capacity

\item
Users able to make upfront payments to reduce their total computing costs even further
\begin{itemize}
	\item
	Standard Reserved Instances (Up to 75\% off on demand)
	
	\item
	Convertible Reserved Instances (Up to 54\% off on demand) capability to change the attributes of the RI as long as the exchange results in the creation of reserved instances of equal or greater value
	
	\item
	Scheduled Reserved Instances are available to launch within the time windows you reserve. This option allows you to match your capacity reservation to a predictable recurring schedule that only requires a fraction of a day, week, or a month.
\end{itemize}
\end{itemize}

\subsection{Spot Instances}
\begin{itemize}
\item
Applications that have flexible start and end times

\item
Application that are only feasible at very low compute prices

\item
Users with urgent computing needs for large amounts of additional capacity
\end{itemize}

\subsection{Dedicated Hosts}
\begin{itemize}
\item
Useful for regulatory requirements that may not support multi tenant virtualization

\item
Great for licensing which does not support multi tenancy or cloud deployments

\item
Can be purchased on Demand (hourly)

\item
Can be purchased as a reservation for u to 70\% off on the on demand price
\end{itemize}

\section{EC2 Instance Types}

\begin{center}
  \begin{tabular}{ c | c | c }
    \hline
    \textbf{Family} & \textbf{Specialty} & \textbf{Use Case} \\ \hline
     D2 & Dense Storage & Fileservers/Data Warehousing/Hadoop \\ \hline
     R4 & Memory Optimized & Memory Intensize Apps/DBs \\ \hline
     M4 & General Purpose & Application Servers \\ \hline
     C4 & Compute Optimized & CPU Intensive Apps/DBs \\ \hline
     G2 & Graphics Intensive & Video Encoding/3D Application Streaming \\ \hline
     I2 & High Speed Storage & NoSQL DBs, Data Warehousing etc \\ \hline
     F1 & Field Programmable Gate Array & Hardware acceleration for your code \\ \hline
     T2 & Low Cost, General Purpose & Web Servers, Small DBs \\ \hline
     P2 & Graphics/General purpose GPU & Machine Learning, Bit Coin Mining etc \\ \hline
     X1 & Memory Optimized & SAP HAN, Apache Spark, etc \\
    
    \hline
  \end{tabular}
\end{center}

\begin{itemize}
\item
D - Density

\item
R - RAM

\item
M - Main Choice for General Purpose Apps

\item
C - Compute

\item
G - Graphics

\item
I - IOPS

\item
F - FPGA

\item
T - Cheap General purpose (think T2 Micro)

\item
P - Graphics (think Pics)

\item
X - Extreme Memory

\end{itemize}

\section{Lab Summary}
\begin{itemize}
\item
Termination Protection is turned off by default, you must turn it on.

\item
On an EBS backed instance, the default action is tfor the root EBS volume to be deleted when the instance is terminated.

\item
EBS root volumes of your default AMIs cannot be encrypted. You can also use a third party tool to encrypt the root volume, or this can be done when creating AMIs in the AWS console or using the API

\item
Additional volumes can be encrypted.

\end{itemize}

\section{Security Group Lab}
\begin{itemize}
\item
All inbound traffic is blocked by default

\item
All outbound traffic is allowed

\item
Changes to security groups take effect immediately

\item
You can have any number of EC2 instances within a security group

\item
Security groups are stateful. If you create an inbound rule allowing traffic in, that traffic is automatically allowed back out again

\item
You can have multiple security groups attached to EC2 instances

\item
You cannot block specific IP addresses using security groups, instead use Network Access Control Lists
\end{itemize}

\section{Volumes and Snapshots}
\begin{itemize}
\item
Volumes exist on EBS - Virtual Hard Disk

\item
Snapshots exist on S3

\item
Snapshots are point in time copies of Volumes

\item
Snapshots are incremental - this means that only the blocks that have changes your last snapshot are moved to S3

\item
If this is your first snapshot, then it will take some time to create

\item
To create a snapshot for Amazon EBS volumes that serve as root devices, you should stop the instance before taking the snapshot. However, you can take a snap while the instance is running

\item
You can create AMIs from EBS-backed instances and snapshots

\item
You can change EBS volume sizes on the fly,including changing the size and storage type

\item
Volumes will always be in the same availability zone as the EC2 instance

\item
To move an EC2 volume from one AZ/region to another, take a snap or an image of it, then copy it to the new AZ/region

\item
Snapshots of encrypted volumes are encrypted automatically

\item
Volumes restored from encrypted snapshots are encrypted automatically

\item
You can share snapshots, but only if they are unencrypted. These snapshots can be shared with other AWS accounts or made public


\end{itemize}

\section{Snapshots of Root Device Volumes}

To create a snapshot for Amazon EBS volumes that serve as root devices, you should stop the instance before taking the snapshot

\section{Volumes vs Snapshots}
\begin{itemize}
\item
Snapshots of encrypted volumes are encrypted automatically

\item
Volumes restored from encrpted snapshots are encrypted automatically

\item
You can share snapshots, but only if they are encrypted. These snapshots can be shared with other AWS accounts or made public

\end{itemize}

\section{AMI Types (EBS vs Instance Store}
You can select you AMI based on:
\begin{itemize}
\item
Region (see regions and availability zones)

\item
Operating System

\item
Architecture (32 bit or 64 bit)

\item
Launch permissions

\item
Storage of root device (Root Device Volume) - Instance Store (Ephemeral Storage) or EBS Backed Volumes
\end{itemize}

All AMIs are categorized as either backed by Amazon EBS or backed by instance store. For EBS volumes, the root device for an instance launched from the AMI is an Amazon EBS volume created from an Amazon EBS snapshot. For instance store volumes, the root device for an instance launched from the AMI is an instance store volume created from a template stored in Amazon S3.

\section{EBS vs Instance Store - Exam Tips}
\begin{itemize}
\item
Instance store volumes are sometimes called ephemeral storage

\item
Instance store volumes cannot be stopped. If the underlying host fails, you will lose your data

\item
EBS backed instances can be stopped. You will not lose the data on this instance if it is stopped

\item
You can reboot both, you will not lose your data

\item
Be default, both root volumes will be deleted on termination, however with EBS volumes, you can tell AWS to keep the root device volume.
\end{itemize}

\section{Elastic Load Balancers}
\begin{itemize}
\item
Instances monitored by ELB are reported as inservice or out of service

\item
Health checks check the instance health by talking to it

\item
Have their own DNS name. You are never given an IP address

\item
Read the ELB FAQ for classic load balancers
\end{itemize}

\section{CloudWatch}
Standard monitoring is always 5 mins. Detailed monitoring is 1 min. \\

What can I do with CloudWatch?
\begin{itemize}
\item
Dashboards - Creates awesome dashboards to see what is happening with our AWS environments

\item
Alarms - Allows you to set Alarms that notify you when particular thesholds are hit

\item
Events - CloudWatch events helps you respond to state changes in your AWS resources

\item
Logs - CloudWatch logs helps you to aggregate, onitor, and store logs
\end{itemize}

\section{Placement Group}
Two Types of placement groups :
\begin{itemize}
\item
Clustered Placement Group

\item
Spread Placement Group
\end{itemize}

\subsection{Clustered Placement Group}
A clustered placement group is a grouping of instances within a single availability zone. Placement groups are recommended for applications that need low network latency, high network throughput, or both. Only certain instances can be launched in to a clustered placement group.

\subsection{Spread Placement Group}
A spread placement group is a group of instances that are places on distinct underlying hardware. Spread placement groups are recommended for applications that have a small number of critical instances that should be kept seperate from each other.

\subsection{Points to remember}
\begin{itemize}
\item
A clustered placement group can't span multiple availability zones.

\item
A spread placement group can

\item
The name you specify for a placement group must be unique within your AWS account

\item
Only certain types of instances can be launched in a placement group (Compute optimized, GPU, Memory optimized, Storage Optimized)

\item
AWS recommend homogeneous instances within placement groups

\item
You cannot merge placement groups

\item
You cannot move an existing instance into a placement group. You can create an AMI from your existing instance, and then launch a new instance from the AMI into a placement group
\end{itemize}

\section{EFS - Elastic File System}
Amazon Elastic File System (EFS) is a file storage service for Amazon Elastic Compute cloud (Amazon EC2) instances. Amazon EFS is easy to use and provides a simple insterface that allows you to create and configure file systems quickly and easily. With Amazon EFS, storage capacity is elastic, growing and shrinking automatically as you add and remove files, so your application have the storage they need, when they need it. \\

Features:
\begin{itemize}
\item
Supports the network file system version 4 (NFSv4) protocol

\item
You only pay for the storage you use (no pre-provisioning required)

\item
Can scale up to the petabytes

\item
Can support thousands of concurrent NFS connections

\item
Data is stored across multiple AZs within a region

\item
Read after write latency

\end{itemize}

\section{AWS Lambda}
AWS Lambda is a compute service where you can upload your code and create a lambda function. AWS lambda take care of provisioning and managing the servers that use to run the code. You don't have to worry about the operating systems, patching, scaling, etc. You can use Lambda in the following ways. As an event driven compute service where AWS lambda runs your code in response to events. These events could be changes to data in an Amazon S3 bucket or an Amazon dynamoDB table. As a compute service to run your code in response to HTTP requests using Amazon aPI gatewat or API calles made using AWS SDKs.

\subsection{How is Lambda priced?}
\begin{itemize}
\item
Number of requests - First 1 million requests are free. \$0.20 per 1 million requests thereafter.

\item
Duration - Duration is calculated from the time your code begins executing until it returns or otherwise terminates, roynded up to the nearest 100ms. The price depends upon the amount of memory you allocate to your function. You are charged \$0.00001667 for every GB-second used. Function cannot execute for more than 5 mins

\subsection{Why is Lambda cool?}
\begin{itemize}
\item
No servers

\item
Continuous scaling

\item
super super super cheap
\end{itemize}

\subsection{Lambda - Exam Tips}
\begin{itemize}
\item
Lambda scales out automatically

\item
Lambda functions are independent, 1 event = 1 function

\item
Lambda is serverless

\item
Know what services are serverless

\item
Lambda functions cna trigger other labda functions, 1 event can = x functions if functions trigger other functions

\item
Architectures can get extremely complicated, AW X-ray allows you to debug what is happening

\item
Lambda can do things globally, you can use to back up S3 buckets to other S3 buckets etc

\item
Know your triggers
\end{itemize}


\end{itemize}
\end{document}